### 2016-06-22 ansi-isolation.tex
Edits.

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt some notation.
An execution trace is destructed into $\A$ and $\visZ$ whenever
individual components of the pair are needed.  Otherwise, it is written
as $\E$. When $\psi$ is a proposition, we write $\underE{\psi}$ to denote
an interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

@@

The ANSI SQL 92 standard requires \iso{Read Committted} isolation to avoid
the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly available
replicated stores. On relational databases, however, {\sc rc} has also come
to be associated with the $\mathtt{MonotonicVis}$ guarantee.
Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ are
sufficient to reason about {\sc rc} isolation on relational stores too. As
we demonstrate in \S\ref{sec:store-consistency}, the combination of
these guarantees with the {\sc sc} property of relational stores
automatically leads to the monotonicity guarantee, which probably
explains why {\sc rc} comes with $\mathtt{MonotonicVis}$ on such stores. On
weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores. \iso{Repeatable
Read} spec ({\sc rr}) extends {\sc mav} with the snapshot visibility guarantee
as described before. \iso{Snapshot Isolation} spec ({\sc si}) extends
{\sc rr} with a one-way serializability guarantee w.r.t the transactions
that perform conflicting writes (i.e., writes to the same shared
variable). Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict\footnote{Fig.~\ref{fig:ansi-isolation} presents slightly
weaker versions of {\sc si} and {\sc ser} specs in the interest of
clarity.}.

>>>>

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt few notations
in this and subsequent sections.  An execution trace is destructed
into $\A$ and $\visZ$ whenever individual components of the pair are
needed. Otherwise, it is written as $\E$. Sometimes, the dot notation
(\eg~$\E.\A$) is also used. Since $\A$ and $\visZ$ are both sets, we
lift the operations on sets to pairs of sets when updating $\E$. For
example, $\E' = \E \cup (\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to
$\E' = (\E.\A \cup \{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.
When $\psi$ is a proposition, $\underE{\psi}$ denotes the
interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

@@

The ANSI SQL 92 standard requires \iso{Read Committted} isolation to
avoid the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which probably
explains why {\sc rc} comes with $\mathtt{MonotonicVis}$ on such
stores. On weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores. Sample
{\sc rc} executions shown in Fig.~\ref{fig:rc-executions} are also
{\sc mav}. \iso{Repeatable Read} spec ({\sc rr}) extends {\sc mav}
with the snapshot visibility guarantee as described before.
\iso{Snapshot Isolation} spec ({\sc si}) extends {\sc rr} with a
one-way serializability guarantee w.r.t the transactions that perform
conflicting writes (i.e., writes to the same shared variable).
Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict\footnote{Fig.~\ref{fig:ansi-isolation} presents slightly
weaker versions of {\sc si} and {\sc ser} specs in the interest of
clarity.}. \GK{ToDo: Figures. Need to figure out a convention to
uniformly represent various isolation levels. Current idea: concurrent
timelines with arrows in both directions.}

### 2016-06-22 store-consistency.tex
Edits.

>>>>

It is worthwhile to note that, although conflict-free $\I_s$ shown
above captures the behaviour of an {\sc SC} database in presence of
isolation constraints, it does not necessarily describe how the
database achieves this behaviour (i.e., its implementation). It is
unusual for databases to check for satisfaction of isolation
constraints everytime an operation is executed. Efficient
implementation strategies (\eg multi-versioning) are employed to
achieve the twin goals of isolation and consistency/recency.
